---
title: "RMD to PPT"
author: "Cheng Peng"
date: "`r Sys.Date()`"
output: powerpoint_presentation
reference_doc: "Galaxy.pptx"
---

```{r setup, include=FALSE}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("pander")) {
   install.packages("pander")
   library(pander)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}
if (!require("patchwork")) {
   install.packages("patchwork")
   library(patchwork)
}
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("gridtext")) {
   install.packages("gridtext")
   library(gridtext)
}

knitr::opts_chunk$set(echo = FALSE)
```


## Auto Industry Indicators

```{r}
# prep frequency table
freqtable <- table(mpg$manufacturer)
df <- as.data.frame.table(freqtable)
theme_set(theme_classic())
# Plot
g <- ggplot(df, aes(Var1, Freq))
g <- g + geom_bar(stat="identity", width = 0.5, fill="tomato2") + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) +
  theme(axis.title.x = element_blank())

t <- richtext_grob("• Hello <br> <br> • Goodbye", halign = 0, valign = -10)
(g+g) / (g+t)
```


## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3


## Slide with R Output

```{r cars, echo = TRUE, comment = NA}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

## Interactive Contents not Allowed

Any R code that generates web-based interactive outcomes will lead to errors! 



